---
title: Test the likelihood calculation integrated across all partiions within a marker
  group for a given subtype
output:
  pdf_document: default
  html_notebook: default
---

# Preliminaries

```{r setwd}
setwd("/Users/chwu/Documents/research/bfc/github/Forensic-Fluids/codes/R/")

```


```{r load_lib}
library(salso)
source("forensicFunctions.R")
load(file = "/Users/chwu/Documents/research/bfc/github/Forensic-Fluids/allPartSets10.RData")
```

```{r load_data}
load("allParts5And7EltsMatsAndSets.rda")

singleTypeFilePath = "data.csv"
single.df = read.csv(singleTypeFilePath, header = TRUE) # whole single sample data
slv.df = single.df[single.df$Sample == 3,] # saliva
dim(slv.df)

sample.df = slv.df[,-1] # single sample data, delete the index column
dim(sample.df)
```

```{r func}
checkAbsDiff = function(threshold = 1e-10, expected = NULL, answer = NULL, print = T){
  test = abs(expected - answer)
  if(print){
    message(paste("Absolute difference", signif(test, 5), sep = ": "))
    message(paste("Test passed", test < threshold, sep = ": "))
  }else{
    return(test < threshold)
  }
  
}

checkAbsRelativeDiff = function(threshold = 1e-10, expected = NULL, answer = NULL, rel = NULL, print = T){
  test = abs(expected - answer)/rel
  if(print){
    message(paste("Absolute relative difference", signif(test, 5), sep = ": "))
    message(paste("Test passed", test < threshold, sep = ": "))
  }else{
    return(test < threshold)
  }
  
}
```

# Partition with three elements

```{r ex1_enum_part}
elts3AllPartMat = enumerate.partitions(3)
# [,1] [,2] [,3]
# [1,]    1    1    1
# [2,]    1    1    2
# [3,]    1    2    1
# [4,]    1    2    2
# [5,]    1    2    3

elts3AllPartSet = apply(elts3AllPartMat, 1, 
                        function(z){
                          setCount = max(z)
                          sets = list()
                          for(setIndex in 1:setCount){
                            sets[[setIndex]] = sort(which(z == setIndex))
                          }
                          
                          return(sets)
                          
                        })

```

```{r ex1_row_sampling}
set.seed(222)
kmeansClust3Index = c(1, 3, 1, 2, 2, 2, 2, 2, 2, 
                      3, 3, 2, 3, 2, 3, 2, 2, 1, 
                      3, 2, 3, 1, 2, 2, 2, 2, 2, 
                      3, 2, 3, 2, 1, 2, 3, 2, 1, 
                      2, 1, 1, 2, 2, 2, 2, 2, 2, 
                      2, 3, 2, 2, 1, 1, 1, 2, 2, 
                      3, 3, 3, 3, 3, 2, 3, 3, 3, 
                      3, 2, 2, 2, 2, 3, 3, 1, 2, 
                      1, 1, 1, 1, 1, 1, 3, 3, 1)
slvClust1ExIndx = sample(which(kmeansClust3Index==1), 2)
# 180 183 
#  75  78
slvClust2ExIndx = sample(which(kmeansClust3Index==2), 5)
# 146 147 145 121 122 
#  41  42  40  16  17 
slvClust3ExIndx = sample(which(kmeansClust3Index==3), 3)
# 185 162 126 
#  80  57  21

rowExIndx = c(slvClust1ExIndx, slvClust2ExIndx, slvClust3ExIndx)
sample.df[rowExIndx ,1:5]
```

```{r ex1_col_sampling}
colExIndx = c(1,2,5)
slvEx1.df = sample.df[rowExIndx, c("CYP", "HBD1", "MUC4")]
slvEx1.df
```

## Test \texttt{calcLikIntAllPart} Function

### Partition 1

Partition: (1, 1, 1)

```{r ex1_part1_clust1}
prod(dim(slvEx1.df))
sum(slvEx1.df)
prod(dim(slvEx1.df)) - sum(slvEx1.df)


```

```{r ex1_part1_calc_lik}
ex1.expt1 = sum(log(1:9)) - sum(log(25:34))
ex1.ans1 = log(calcLikIntAllPart(eltsAllPartSet = list(list(c(1, 2, 3))), 
                             sample = slvEx1.df, 
                             alphaC = 2, 
                             betaC = 3))

checkAbsDiff(expected = ex1.expt1, 
             answer = ex1.ans1)
checkAbsRelativeDiff(expected = ex1.expt1, 
                     answer = ex1.ans1, 
                     rel = min(abs(c(ex1.expt1, ex1.ans1))))
```

```{r ex1_part1_calc_mdp}
ex1.mdp.expt1 = 15/(7 * 13)
ex1.mdp.ans1 = calcMDPColPrior(alpha = 12, setList = list(c(1,1,1)), setCountMax = 3)

checkAbsDiff(expected = ex1.mdp.expt1, 
             answer = ex1.mdp.ans1)
checkAbsRelativeDiff(expected = ex1.mdp.expt1, 
                     answer = ex1.mdp.ans1, 
                     rel = min(abs(c(ex1.mdp.expt1, ex1.mdp.ans1))))
```

### Partition 2

Partition: (1, 1, 2)

```{r ex1_part2_clust1}
prod(dim(slvEx1.df[, 1:2]))
sum(slvEx1.df[, 1:2])
prod(dim(slvEx1.df[, 1:2])) - sum(slvEx1.df[, 1:2])
```

```{r ex1_part2_clust2}
length(slvEx1.df[, 3])
sum(slvEx1.df[, 3])
```

```{r ex1_part2_calc_lik}
ex1.expt2 = (sum(log(1:4)) - sum(log(20:24))) + (sum(log(1:6)) - sum(log(8:14)))

ex1.ans2 = log(calcLikIntAllPart(eltsAllPartSet = list(list(c(1, 2), c(3))), 
                  sample = slvEx1.df, 
                  alphaC = 2, 
                  betaC = 3))

checkAbsDiff(expected = ex1.expt2, 
             answer = ex1.ans2)
checkAbsRelativeDiff(expected = ex1.expt2, 
                     answer = ex1.ans2, 
                     rel = min(abs(c(ex1.expt2, ex1.ans2))))
```

```{r ex1_part2_calc_mdp}
ex1.mdp.expt2 = 20/(7 * 13)
ex1.mdp.ans2 = calcMDPColPrior(alpha = 12, setList = list(c(1,1), c(3)), setCountMax = 3)

abs(ex1.mdp.expt2 - ex1.mdp.ans2)/max(c(ex1.mdp.expt2, ex1.mdp.ans2))
abs((ex1.mdp.expt2 - ex1.mdp.ans2)/abs(max(c(ex1.mdp.expt2, ex1.mdp.ans2)))) < 1e-10

checkAbsDiff(expected = ex1.mdp.expt2, 
             answer = ex1.mdp.ans2)
checkAbsRelativeDiff(expected = ex1.mdp.expt2, 
                     answer = ex1.mdp.ans2, 
                     rel = min(abs(c(ex1.mdp.expt2, ex1.mdp.ans2))))
```

### Partition 3

Partition: (1, 2, 1)

```{r ex1_part3_clust1}
prod(dim(slvEx1.df[, c(1,3)]))
sum(slvEx1.df[, c(1,3)])
```

```{r ex1_part3_clust2}
length(slvEx1.df[, 2]) - sum(slvEx1.df[, 2])
```

```{r ex1_part3_calc_lik}
ex1.expt3 = ((sum(log(1:7)) - sum(log(17:24)))) + 
  ((sum(log(1:3)) - sum(log(11:14))))
ex1.ans3 = log(calcLikIntAllPart(eltsAllPartSet = list(list(c(1, 3), c(2))), 
                  sample = slvEx1.df, 
                  alphaC = 2, 
                  betaC = 3))

checkAbsDiff(expected = ex1.expt3, 
             answer = ex1.ans3)
checkAbsRelativeDiff(expected = ex1.expt3, 
                     answer = ex1.ans3, 
                     rel = min(abs(c(ex1.expt3, ex1.ans3))))
```

```{r ex1_part3_calc_mdp}
ex1.mdp.expt3 = 20/(7 * 13)
ex1.mdp.ans3 = calcMDPColPrior(alpha = 12, 
                                  setList = list(c(1,3), c(2)), 
                                  setCountMax = 3)

checkAbsDiff(expected = ex1.mdp.expt3, 
             answer = ex1.mdp.ans3)
checkAbsRelativeDiff(expected = ex1.mdp.expt3, 
                     answer = ex1.mdp.ans3, 
                     rel = min(abs(c(ex1.mdp.expt3, ex1.mdp.ans3))))
```

### Partition 4

Partition: (1, 2, 2)

```{r ex1_part4_clust1}
# 1    2    2
length(slvEx1.df[, 1])
sum(slvEx1.df[, 1])
```

```{r ex1_part4_clust2}
prod(dim(slvEx1.df[, c(2,3)]))
sum(slvEx1.df[, c(2,3)])
```

```{r ex1_part4_calc_lik}
ex1.expt4 = -sum(log(c(6, 13, 14))) + 
  ((sum(log(1:8)) - sum(log(16:24))))
ex1.ans4 = log(calcLikIntAllPart(eltsAllPartSet = list(list(c(1), c(2, 3))), 
                         sample = slvEx1.df, 
                         alphaC = 2, 
                         betaC = 3))

checkAbsDiff(expected = ex1.expt4, 
             answer = ex1.ans4)
checkAbsRelativeDiff(expected = ex1.expt4, 
                     answer = ex1.ans4, 
                     rel = min(abs(c(ex1.expt4, ex1.ans4))))
```

```{r ex1_part4_calc_mdp}
ex1.mdp.expt4 = 20/(7 * 13)
ex1.mdp.ans4 = calcMDPColPrior(alpha = 12, 
                                  setList = list(c(2,3), c(1)), 
                                  setCountMax = 3)

checkAbsDiff(expected = ex1.mdp.expt4, 
             answer = ex1.mdp.ans4)
checkAbsRelativeDiff(expected = ex1.mdp.expt4, 
                     answer = ex1.mdp.ans4, 
                     rel = min(abs(c(ex1.mdp.expt4, ex1.mdp.ans4))))
```


### Partition 5

Partition: (1, 2, 3)

```{r ex1_part5_clust1}
length(slvEx1.df[, 1])
sum(slvEx1.df[, 1])
```

```{r ex1_part5_clust2}
length(slvEx1.df[, 2])
sum(slvEx1.df[, 2])
```

```{r ex1_part5_clust3}
length(slvEx1.df[, 3])
sum(slvEx1.df[, 3])
```

```{r ex1_part5_calc_lik}
ex1.expt5 = (-sum(log(c(6, 13, 14)))) + 
  -sum(log(c(2, 11, 13, 14))) + 
  (sum(log(1:6)) - sum(log(c(8:14))))
ex1.ans5 = log(calcLikIntAllPart(eltsAllPartSet = list(list(c(1), c(2), c(3))), 
                         sample = slvEx1.df, 
                         alphaC = 2, 
                         betaC = 3))

checkAbsDiff(expected = ex1.expt5, 
             answer = ex1.ans5)
checkAbsRelativeDiff(expected = ex1.expt5,
                     answer = ex1.ans5, 
                     rel = min(abs(c(ex1.expt5, ex1.ans5))))
```


```{r ex1_part5_calc_mdp}
ex1.mdp.expt5 = 16/(7 * 13)
ex1.mdp.ans5 = calcMDPColPrior(alpha = 12, 
                                  setList = list(c(1), c(2), c(3)), 
                                  setCountMax = 3)

checkAbsDiff(expected = ex1.mdp.expt5, 
             answer = ex1.mdp.ans5)
checkAbsRelativeDiff(expected = ex1.mdp.expt5,
                     answer = ex1.mdp.ans5, 
                     rel = min(abs(c(ex1.mdp.expt5, ex1.mdp.ans5))))
```

## Test \texttt{calcLogSubTypeLik} Function

```{r ex1_subtype_lik}
ex1.subtype.lik.expt = ex1.mdp.expt1*exp(ex1.expt1) +
  ex1.mdp.expt2*exp(ex1.expt2) +
  ex1.mdp.expt3*exp(ex1.expt3) +
  ex1.mdp.expt4*exp(ex1.expt4) +
  ex1.mdp.expt5*exp(ex1.expt5)

ex1.subtype.log.lik.expt = log(ex1.subtype.lik.expt)

elts3AllPartSetMDPPrior = unlist(lapply(elts3AllPartSet, 
                                        calcMDPColPrior, 
                                        alpha = 12, 
                                        setCountMax = 3))
ex1.subtype.log.lik.ans = calcLogSubTypeLik(eltsAllPartSetList = list(elts3AllPartSet), 
                                            eltsAllPartSetPriorList = list(elts3AllPartSetMDPPrior),
                                            sample = slvEx1.df, 
                                            mkrList = list(colnames(slvEx1.df)), 
                                            alphaC = 2, 
                                            betaC = 3)

checkAbsDiff(expected = ex1.subtype.log.lik.expt, 
             answer = ex1.subtype.log.lik.ans)
checkAbsRelativeDiff(expected = ex1.subtype.log.lik.expt,
                     answer = ex1.subtype.log.lik.ans, 
                     rel = min(abs(c(ex1.subtype.log.lik.expt, 
                                     ex1.subtype.log.lik.ans))))
```

# Partition with four elements

```{r ex2_enum_part}
elts4AllPartMat = enumerate.partitions(4)
#       [,1] [,2] [,3] [,4]
#  [1,]    1    1    1    1
#  [2,]    1    1    1    2
#  [3,]    1    1    2    1
#  [4,]    1    1    2    2
#  [5,]    1    1    2    3
#  [6,]    1    2    1    1
#  [7,]    1    2    1    2
#  [8,]    1    2    1    3
#  [9,]    1    2    2    1
# [10,]    1    2    2    2
# [11,]    1    2    2    3
# [12,]    1    2    3    1
# [13,]    1    2    3    2
# [14,]    1    2    3    3
# [15,]    1    2    3    4

elts4AllPartSet = apply(elts4AllPartMat, 1, 
                        function(z){
                          setCount = max(z)
                          sets = list()
                          for(setIndex in 1:setCount){
                            sets[[setIndex]] = sort(which(z == setIndex))
                          }
                          
                          return(sets)
                          
                        })
```



```{r}
set.seed(123)
slvClust1ExIndx2 = sample(which(kmeansClust3Index==1), 2)
# 180 179 
#  75  74 
slvClust2ExIndx2 = sample(which(kmeansClust3Index==2), 5)
# 111 173 130 148 149 
#   6  68  25  43  44
slvClust3ExIndx2 = sample(which(kmeansClust3Index==3), 3)
# 120 168 135 
#  15  63  30 
```

```{r}
rowExIndx2 = c(slvClust1ExIndx2, slvClust2ExIndx2, slvClust3ExIndx2)
slvEx2.df = sample.df[rowExIndx2 ,c("HTN3", "MUC7", "PRB4", "SMR3B" )]
slvEx2.df
```

## Test \texttt{} Function

### Partition 1

```{r ex2_part1_clust1}
sum(slvEx2.df)
prod(dim(slvEx2.df))
```


```{r ex2_part1_calc_lik}
ex2.expt1 = sum(log(1:13)) - sum(log(31:44))
ex2.ans1 = log(calcLikIntAllPart(eltsAllPartSet = list(list(c(1:4))), 
                         sample = slvEx2.df, 
                         alphaC = 3, 
                         betaC = 2))

checkAbsDiff(expected = ex2.expt1, 
             answer = ex2.ans1)
checkAbsRelativeDiff(expected = ex2.expt1,
                     answer = ex2.ans1, 
                     rel = min(abs(c(ex2.expt1, ex2.ans1))))
```


```{r ex2_part1_calc_mdp}
ex2.mdp.expt1 = 35/(3 * 17 * 19)
ex2.mdp.ans1 = calcMDPColPrior(alpha = 16, 
                                  setList = list(c(1, 2, 3, 4)), 
                                  setCountMax = 4)

checkAbsDiff(expected = ex2.mdp.expt1, 
             answer = ex2.mdp.ans1)
checkAbsRelativeDiff(expected = ex2.mdp.expt1,
                     answer = ex2.mdp.ans1, 
                     rel = min(abs(c(ex2.mdp.expt1, ex2.mdp.ans1))))
```

### Partition 2

Partition pattern: (1, 1, 1, 2)

```{r ex2_part2_clust1}
sum(slvEx2.df[, 1:3])
prod(dim(slvEx2.df[, 1:3]))
```


```{r ex2_part2_clust2}
sum(slvEx2.df[,4])
length(slvEx2.df[,4])
```

```{r ex2_part2_calc_lik}
ex2.expt2.clust1.log.lik = sum(log(1:12)) - sum(log(22:34))
ex2.expt2.clust2.log.lik = -(log(6) + log(13) + log(14))

ex2.expt2 = ex2.expt2.clust1.log.lik + ex2.expt2.clust2.log.lik
ex2.ans2 = log(calcLikIntAllPart(eltsAllPartSet = list(list(c(1:3), c(4))), 
                         sample = slvEx2.df, 
                         alphaC = 3, 
                         betaC = 2))

checkAbsDiff(expected = ex2.expt2, 
             answer = ex2.ans2)
checkAbsRelativeDiff(expected = ex2.expt2,
                     answer = ex2.ans2, 
                     rel = min(abs(c(ex2.expt2, ex2.ans2))))
```


```{r ex2_part2_calc_mdp}
ex2.mdp.expt2 = 20/(17 * 19)
ex2.mdp.ans2 = calcMDPColPrior(alpha = 16, 
                                  setList = list(c(1, 2, 3), c(4)), 
                                  setCountMax = 4)

checkAbsDiff(expected = ex2.mdp.expt2, 
             answer = ex2.mdp.ans2)
checkAbsRelativeDiff(expected = ex2.mdp.expt2,
                     answer = ex2.mdp.ans2, 
                     rel = min(abs(c(ex2.mdp.expt2, ex2.mdp.ans2))))
```

### Partition 3

Partition pattern: (1, 1, 2, 1)

```{r ex2_part3_clust1}
sum(slvEx2.df[, c(1, 2, 4)])
prod(dim(slvEx2.df[, c(1, 2, 4)]))
```


```{r ex2_part3_clust2}
sum(slvEx2.df[,3])
length(slvEx2.df[,3])
```

```{r ex2_part3_calc_lik}
ex2.expt3.clust1.log.lik = sum(log(1:5)) - sum(log(29:34))
ex2.expt3.clust2.log.lik = -(log(5) + log(11) + log(13) + log(14))

ex2.expt3 = ex2.expt3.clust1.log.lik + ex2.expt3.clust2.log.lik
ex2.ans3 = log(calcLikIntAllPart(eltsAllPartSet = list(list(c(1,2,4), c(3))), 
                         sample = slvEx2.df, 
                         alphaC = 3, 
                         betaC = 2))

checkAbsDiff(expected = ex2.expt3, 
             answer = ex2.ans3)
checkAbsRelativeDiff(expected = ex2.expt3,
                     answer = ex2.ans3, 
                     rel = min(abs(c(ex2.expt3, ex2.ans3))))
```

```{r ex2_part3_calc_mdp}
ex2.mdp.expt3 = 20/(17 * 19)
ex2.mdp.ans3 = calcMDPColPrior(alpha = 16, 
                                  setList = list(c(1, 2, 4), c(3)), 
                                  setCountMax = 4)

checkAbsDiff(expected = ex2.mdp.expt3, 
             answer = ex2.mdp.ans3)
checkAbsRelativeDiff(expected = ex2.mdp.expt3,
                     answer = ex2.mdp.ans3, 
                     rel = min(abs(c(ex2.mdp.expt3, ex2.mdp.ans3))))
```

### Partition 4

Partition pattern: (1, 1, 2, 2)

```{r ex2_part4_clust1}
sum(slvEx2.df[, c(1, 2)])
prod(dim(slvEx2.df[, c(1, 2)]))
```


```{r ex2_part4_clust2}
sum(slvEx2.df[, c(3, 4)])
prod(dim(slvEx2.df[, c(3, 4)]))
```

```{r ex2_part4_calc_lik}
ex2.expt4.clust1.log.lik = - sum(log(20:23))
ex2.expt4.clust2.log.lik = sum(log(1:10)) - sum(log(14:24))

ex2.expt4 = ex2.expt4.clust1.log.lik + ex2.expt4.clust2.log.lik
ex2.ans4 = log(calcLikIntAllPart(eltsAllPartSet = list(list(c(1,2), c(3,4))), 
                         sample = slvEx2.df, 
                         alphaC = 3, 
                         betaC = 2))

checkAbsDiff(expected = ex2.expt4, 
             answer = ex2.ans4)
checkAbsRelativeDiff(expected = ex2.expt4,
                     answer = ex2.ans4, 
                     rel = min(abs(c(ex2.expt4, ex2.ans4))))
```


```{r ex2_part4_calc_mdp}
ex2.mdp.expt4 = 50/(3 * 17 * 19)
ex2.mdp.ans4 = calcMDPColPrior(alpha = 16, 
                                  setList = list(c(1, 2), c(3, 4)), 
                                  setCountMax = 4)

checkAbsDiff(expected = ex2.mdp.expt4, 
             answer = ex2.mdp.ans4)
checkAbsRelativeDiff(expected = ex2.mdp.expt4,
                     answer = ex2.mdp.ans4, 
                     rel = min(abs(c(ex2.mdp.expt4, ex2.mdp.ans4))))
```

### Partition 5

Partition pattern: (1, 1, 2, 3)

```{r ex2_part5_clust1}
sum(slvEx2.df[, c(1, 2)])
prod(dim(slvEx2.df[, c(1, 2)]))
```


```{r ex2_part5_clust2}
sum(slvEx2.df[, 3])
length(slvEx2.df[, 3])
```

```{r ex2_part5_clust3}
sum(slvEx2.df[, 4])
length(slvEx2.df[,4])
```

```{r ex2_part5_calc_lik}
ex2.expt5.clust1.log.lik = - sum(log(20:23))
ex2.expt5.clust2.log.lik = - sum(log(c(5, 11, 13, 14)))
ex2.expt5.clust3.log.lik = - sum(log(c(6, 13, 14)))

ex2.expt5 = ex2.expt5.clust1.log.lik + ex2.expt5.clust2.log.lik + ex2.expt5.clust3.log.lik
ex2.ans5 = log(calcLikIntAllPart(eltsAllPartSet = list(list(c(1,2), c(3), c(4))), 
                         sample = slvEx2.df, 
                         alphaC = 3, 
                         betaC = 2))

checkAbsDiff(expected = ex2.expt5, 
             answer = ex2.ans5)
checkAbsRelativeDiff(expected = ex2.expt5,
                     answer = ex2.ans5, 
                     rel = min(abs(c(ex2.expt5, ex2.ans5))))
```


```{r ex2_part5_calc_mdp}
ex2.mdp.expt5 = 80/(3 * 17 * 19)
ex2.mdp.ans5 = calcMDPColPrior(alpha = 16, 
                               setList = list(c(1, 2), c(3), c(4)), 
                               setCountMax = 4)

checkAbsDiff(expected = ex2.mdp.expt5, 
             answer = ex2.mdp.ans5)
checkAbsRelativeDiff(expected = ex2.mdp.expt5,
                     answer = ex2.mdp.ans5, 
                     rel = min(abs(c(ex2.mdp.expt5, ex2.mdp.ans5))))
```

### Partition 6

Partition pattern: (1, 2, 1, 1)

```{r ex2_part6_clust1}
sum(slvEx2.df[, c(1, 3, 4)])
prod(dim(slvEx2.df[, c(1, 3, 4)]))
```


```{r ex2_part6_clust2}
sum(slvEx2.df[, 2])
length(slvEx2.df[, 2])
```

```{r ex2_part6_calc_lik}
ex2.expt6.clust1.log.lik = sum(log(1:13)) - sum(log(21:34))
ex2.expt6.clust2.log.lik = -(log(13) + log(14))

ex2.expt6 = ex2.expt6.clust1.log.lik + ex2.expt6.clust2.log.lik
ex2.ans6 = log(calcLikIntAllPart(eltsAllPartSet = list(list(c(1,3,4), c(2))), 
                         sample = slvEx2.df, 
                         alphaC = 3, 
                         betaC = 2))

checkAbsDiff(expected = ex2.expt6, 
             answer = ex2.ans6)
checkAbsRelativeDiff(expected = ex2.expt6,
                     answer = ex2.ans6, 
                     rel = min(abs(c(ex2.expt6, ex2.ans6))))
```


```{r ex2_part6_calc_mdp}
ex2.mdp.expt6 = 20/(17 * 19)
ex2.mdp.ans6 = calcMDPColPrior(alpha = 16, 
                               setList = list(c(1, 3, 4), c(2)), 
                               setCountMax = 4)

checkAbsDiff(expected = ex2.mdp.expt6, 
             answer = ex2.mdp.ans6)
checkAbsRelativeDiff(expected = ex2.mdp.expt6,
                     answer = ex2.mdp.ans6, 
                     rel = min(abs(c(ex2.mdp.expt6, ex2.mdp.ans6))))
```

### Partition 7

Partition: (1, 2, 1, 2)


```{r ex2_part7_clust1}
sum(slvEx2.df[, c(1, 3)])
prod(dim(slvEx2.df[, c(1, 3)]))
```


```{r ex2_part7_clust2}
sum(slvEx2.df[, c(2, 4)])
prod(dim(slvEx2.df[, c(2, 4)]))
```

```{r ex2_part7_calc_lik}
ex2.expt7.clust1.log.lik = sum(log(1:11)) - sum(log(13:24))
ex2.expt7.clust2.log.lik = -sum(log(c(11, 23, 24)))

ex2.expt7 = ex2.expt7.clust1.log.lik + ex2.expt7.clust2.log.lik
ex2.ans7 = log(calcLikIntAllPart(eltsAllPartSet = list(list(c(1,3), c(2,4))), 
                         sample = slvEx2.df, 
                         alphaC = 3, 
                         betaC = 2))

checkAbsDiff(expected = ex2.expt7, 
             answer = ex2.ans7)
checkAbsRelativeDiff(expected = ex2.expt7,
                     answer = ex2.ans7, 
                     rel = min(abs(c(ex2.expt7, ex2.ans7))))
```


```{r ex2_part7_calc_mdp}
ex2.mdp.expt7 = 50/(3 * 17 * 19)
ex2.mdp.ans7 = calcMDPColPrior(alpha = 16, 
                               setList = list(c(1, 3), c(2, 4)), 
                               setCountMax = 4)

checkAbsDiff(expected = ex2.mdp.expt7, 
             answer = ex2.mdp.ans7)
checkAbsRelativeDiff(expected = ex2.mdp.expt7,
                     answer = ex2.mdp.ans7, 
                     rel = min(abs(c(ex2.mdp.expt7, ex2.mdp.ans7))))
```

### Partition 8

Partition pattern: (1, 2, 1, 3)

```{r ex2_part8_clust1}
sum(slvEx2.df[, c(1, 3)])
prod(dim(slvEx2.df[, c(1, 3)]))
```


```{r ex2_part8_clust2}
sum(slvEx2.df[, 2])
length(slvEx2.df[, 2])
```

```{r ex2_part8_clust3}
sum(slvEx2.df[, 4])
length(slvEx2.df[,4])
```

```{r ex2_part8_calc_lik}
ex2.expt8.clust1.log.lik = sum(log(1:11)) - sum(log(13:24))
ex2.expt8.clust2.log.lik = - (log(13) + log(14))
ex2.expt8.clust3.log.lik = - sum(log(c(6, 13, 14)))

ex2.expt8 = ex2.expt8.clust1.log.lik + ex2.expt8.clust2.log.lik + ex2.expt8.clust3.log.lik
ex2.ans8 = log(calcLikIntAllPart(eltsAllPartSet = list(list(c(1,3), c(2), c(4))), 
                         sample = slvEx2.df, 
                         alphaC = 3, 
                         betaC = 2))

checkAbsDiff(expected = ex2.expt8, 
             answer = ex2.ans8)
checkAbsRelativeDiff(expected = ex2.expt8,
                     answer = ex2.ans8, 
                     rel = min(abs(c(ex2.expt8, ex2.ans8))))
```


```{r ex2_part8_calc_mdp}
ex2.mdp.expt8 = 80/(3 * 17 * 19)
ex2.mdp.ans8 = calcMDPColPrior(alpha = 16, 
                               setList = list(c(1, 3), c(2), c(4)), 
                               setCountMax = 4)

checkAbsDiff(expected = ex2.mdp.expt8, 
             answer = ex2.mdp.ans8)
checkAbsRelativeDiff(expected = ex2.mdp.expt8,
                     answer = ex2.mdp.ans8,
                     rel = min(abs(c(ex2.mdp.expt8, ex2.mdp.ans8))))
```

### Partition 9

Partition: (1, 2, 2, 1)


```{r ex2_part9_clust1}
sum(slvEx2.df[, c(1, 4)])
prod(dim(slvEx2.df[, c(1, 4)]))
```


```{r ex2_part9_clust2}
sum(slvEx2.df[, c(2, 3)])
prod(dim(slvEx2.df[, c(2, 3)]))
```

```{r ex2_part9_calc_lik}
ex2.expt9.clust1.log.lik = sum(log(1:5)) - sum(log(19:24))
ex2.expt9.clust2.log.lik = sum(log(1:9)) - sum(log(15:24))

ex2.expt9 = ex2.expt9.clust1.log.lik + ex2.expt9.clust2.log.lik
ex2.ans9 = log(calcLikIntAllPart(eltsAllPartSet = list(list(c(1,4), c(2,3))), 
                         sample = slvEx2.df, 
                         alphaC = 3, 
                         betaC = 2))

checkAbsDiff(expected = ex2.expt9, 
             answer = ex2.ans9)
checkAbsRelativeDiff(expected = ex2.expt9,
                     answer = ex2.ans9, 
                     rel = min(abs(c(ex2.expt9, ex2.ans9))))
```



```{r ex2_part9_calc_mdp}
ex2.mdp.expt9 = 50/(3 * 17 * 19)
ex2.mdp.ans9 = calcMDPColPrior(alpha = 16, 
                               setList = list(c(1, 4), c(2, 3)), 
                               setCountMax = 4)

checkAbsDiff(expected = ex2.mdp.expt9, 
             answer = ex2.mdp.ans9)
checkAbsRelativeDiff(expected = ex2.mdp.expt9,
                     answer = ex2.mdp.ans9, 
                     rel = min(abs(c(ex2.mdp.expt9, ex2.mdp.ans9))))
```

### Partition 10

Partition: (1, 2, 2, 2)



```{r ex2_part10_clust1}
sum(slvEx2.df[, 1])
length(slvEx2.df[, 1])
```


```{r ex2_part10_clust2}
sum(slvEx2.df[, 2:4])
prod(dim(slvEx2.df[, 2:4]))
```

```{r ex2_part10_calc_lik}
ex2.expt10.clust1.log.lik = -sum(log(5) + log(11) + log(13) + log(14))
ex2.expt10.clust2.log.lik = sum(log(1:10)) - sum(log(24:34))

ex2.expt10 = ex2.expt10.clust1.log.lik + ex2.expt10.clust2.log.lik
ex2.ans10 = log(calcLikIntAllPart(eltsAllPartSet = list(list(c(1), c(2:4))), 
                         sample = slvEx2.df, 
                         alphaC = 3, 
                         betaC = 2))

checkAbsDiff(expected = ex2.expt10, 
             answer = ex2.ans10)
checkAbsRelativeDiff(expected = ex2.expt10,
                     answer = ex2.ans10, 
                     rel = min(abs(c(ex2.expt10, ex2.ans10))))
```

```{r ex2_part10_calc_mdp}
ex2.mdp.expt10 = 20/(17 * 19)
ex2.mdp.ans10 = calcMDPColPrior(alpha = 16, 
                                setList = list(c(1), c(2, 3, 4)), 
                                setCountMax = 4)

checkAbsDiff(expected = ex2.mdp.expt10, 
             answer = ex2.mdp.ans10)
checkAbsRelativeDiff(expected = ex2.mdp.expt10,
                     answer = ex2.mdp.ans10, 
                     rel = min(abs(c(ex2.mdp.expt10, ex2.mdp.ans10))))
```

### Partition 11

Partition: (1, 2, 2, 3)

```{r ex2_part11_clust1}
sum(slvEx2.df[, 1])
length(slvEx2.df[, 1])
```


```{r ex2_part11_clust2}
sum(slvEx2.df[, c(2, 3)])
prod(dim(slvEx2.df[, c(2, 3)]))

```

```{r ex2_part11_clust3}
sum(slvEx2.df[, 4])
length(slvEx2.df[,4])
```

```{r ex2_part11_calc_lik}
ex2.expt11.clust1.log.lik = -sum(log(c(5, 11, 13, 14)))
ex2.expt11.clust2.log.lik = sum(log(1:9)) - sum(log(15:24))
ex2.expt11.clust3.log.lik = -sum(log(c(6, 13, 14)))

ex2.expt11 = ex2.expt11.clust1.log.lik + 
  ex2.expt11.clust2.log.lik + 
  ex2.expt11.clust3.log.lik
ex2.ans11 = log(calcLikIntAllPart(eltsAllPartSet = list(list(c(1), c(2, 3), c(4))), 
                         sample = slvEx2.df, 
                         alphaC = 3, 
                         betaC = 2))

checkAbsDiff(expected = ex2.expt11, 
             answer = ex2.ans11)
checkAbsRelativeDiff(expected = ex2.expt11,
                     answer = ex2.ans11, 
                     rel = min(abs(c(ex2.expt11, ex2.ans11))))
```



```{r ex2_part11_calc_mdp}
ex2.mdp.expt11 = 80/(3 * 17 * 19)
ex2.mdp.ans11 = calcMDPColPrior(alpha = 16, 
                                setList = list(c(1), c(2, 3), c(4)), 
                                setCountMax = 4)

checkAbsDiff(expected = ex2.mdp.expt11, 
             answer = ex2.mdp.ans11)
checkAbsRelativeDiff(expected = ex2.mdp.expt11,
                     answer = ex2.mdp.ans11, 
                     rel = min(abs(c(ex2.mdp.expt11, ex2.mdp.ans11))))
```

### Partition 12

Partition: (1, 2, 3, 1)

```{r ex2_part12_clust1}
sum(slvEx2.df[, c(1, 4)])
prod(dim(slvEx2.df[, c(1, 4)]))
```


```{r ex2_part12_clust2}
sum(slvEx2.df[, 2])
length(slvEx2.df[, 2])
```

```{r ex2_part12_clust3}
sum(slvEx2.df[, 3])
length(slvEx2.df[, 3])
```

```{r ex2_part12_calc_lik}
ex2.expt12.clust1.log.lik = sum(log(1:5)) - sum(log(19:24))
ex2.expt12.clust2.log.lik = - sum(log(13) + log(14))
ex2.expt12.clust3.log.lik = -sum(log(c(5, 11, 13, 14)))

ex2.expt12 = ex2.expt12.clust1.log.lik + 
  ex2.expt12.clust2.log.lik + 
  ex2.expt12.clust3.log.lik
ex2.ans12 = log(calcLikIntAllPart(eltsAllPartSet = list(list(c(1, 4), c(2), c(3))), 
                         sample = slvEx2.df, 
                         alphaC = 3, 
                         betaC = 2))

checkAbsDiff(expected = ex2.expt12, 
             answer = ex2.ans12)
checkAbsRelativeDiff(expected = ex2.expt12,
                     answer = ex2.ans12, 
                     rel = min(abs(c(ex2.expt12, ex2.ans12))))
```



```{r ex2_part12_calc_mdp}
ex2.mdp.expt12 = 80/(3 * 17 * 19)
ex2.mdp.ans12 = calcMDPColPrior(alpha = 16, 
                                setList = list(c(1, 4), c(2), c(3)), 
                                setCountMax = 4)

checkAbsDiff(expected = ex2.mdp.expt12,
             answer = ex2.mdp.ans12)
checkAbsRelativeDiff(expected = ex2.mdp.expt12,
                     answer = ex2.mdp.ans12, 
                     rel = min(abs(c(ex2.mdp.expt12, ex2.mdp.ans12))))
```

### Partition 13

Partition: (1, 2, 3, 2)


```{r ex2_part13_clust1}
sum(slvEx2.df[, 1])
length(slvEx2.df[, 1])
```


```{r ex2_part13_clust2}

sum(slvEx2.df[, c(2, 4)])
prod(dim(slvEx2.df[, c(2, 4)]))
```

```{r ex2_part13_clust3}
sum(slvEx2.df[, 3])
length(slvEx2.df[, 3])
```

```{r ex2_part13_calc_lik}
ex2.expt13.clust1.log.lik = -sum(log(c(5, 11, 13, 14)))
ex2.expt13.clust2.log.lik = - (log(11) + log(23) + log(24))
ex2.expt13.clust3.log.lik = -sum(log(c(5, 11, 13, 14)))

ex2.expt13 = ex2.expt13.clust1.log.lik + 
  ex2.expt13.clust2.log.lik + 
  ex2.expt13.clust3.log.lik
ex2.ans13 = log(calcLikIntAllPart(eltsAllPartSet = list(list(c(1), c(2, 4), c(3))), 
                         sample = slvEx2.df, 
                         alphaC = 3, 
                         betaC = 2))

checkAbsDiff(expected = ex2.expt13, 
             answer = ex2.ans13)
checkAbsRelativeDiff(expected = ex2.expt13,
                     answer = ex2.ans13,
                     rel = min(abs(c(ex2.expt13, ex2.ans13))))
```


```{r ex2_part13_calc_mdp}
ex2.mdp.expt13 = 80/(3 * 17 * 19)
ex2.mdp.ans13 = calcMDPColPrior(alpha = 16, 
                                setList = list(c(1), c(2, 4), c(3)), 
                                setCountMax = 4)

checkAbsDiff(expected = ex2.mdp.expt13,
             answer = ex2.mdp.ans13)
checkAbsRelativeDiff(expected = ex2.mdp.expt13,
                     answer = ex2.mdp.ans13, 
                     rel = min(abs(c(ex2.mdp.expt13, ex2.mdp.ans13))))
```

### Partition 14

Partition: (1, 2, 3, 3)

```{r ex2_part14_clust1}
sum(slvEx2.df[, 1])
length(slvEx2.df[, 1])
```


```{r ex2_part14_clust2}
sum(slvEx2.df[, 2])
length(slvEx2.df[, 2])
```

```{r ex2_part14_clust3}
sum(slvEx2.df[, c(3, 4)])
prod(dim(slvEx2.df[, c(3, 4)]))
```

```{r ex2_part14_calc_lik}
ex2.expt14.clust1.log.lik = -sum(log(c(5, 11, 13, 14)))
ex2.expt14.clust2.log.lik = - (log(13) + log(14))
ex2.expt14.clust3.log.lik = sum(log(1:10)) -sum(log(c(14:24)))

ex2.expt14 = ex2.expt14.clust1.log.lik + 
  ex2.expt14.clust2.log.lik + 
  ex2.expt14.clust3.log.lik
ex2.ans14 = log(calcLikIntAllPart(eltsAllPartSet = list(list(c(1), c(2), c(3, 4))), 
                         sample = slvEx2.df, 
                         alphaC = 3, 
                         betaC = 2))

checkAbsDiff(expected = ex2.expt14, 
             answer = ex2.ans14)
checkAbsRelativeDiff(expected = ex2.expt14,
                     answer = ex2.ans14,
                     rel = min(abs(c(ex2.expt14, ex2.ans14))))
```



```{r ex2_part14_calc_mdp}
ex2.mdp.expt14 = 80/(3 * 17 * 19)
ex2.mdp.ans14 = calcMDPColPrior(alpha = 16, 
                                setList = list(c(1), c(2), c(3, 4)), 
                                setCountMax = 4)

checkAbsDiff(expected = ex2.mdp.expt14,
             answer = ex2.mdp.ans14)
checkAbsRelativeDiff(expected = ex2.mdp.expt14,
                     answer = ex2.mdp.ans14, 
                     rel = min(abs(c(ex2.mdp.expt14, ex2.mdp.ans14))))
```

### Partition 15


Partition: (1, 2, 3, 4)

```{r ex2_part15_clust1}
sum(slvEx2.df[, 1])
length(slvEx2.df[, 1])
```


```{r ex2_part15_clust2}
sum(slvEx2.df[, 2])
length(slvEx2.df[, 2])
```

```{r ex2_part15_clust3}
sum(slvEx2.df[, 3])
length(slvEx2.df[, 3])
```

```{r ex2_part15_clust4}
sum(slvEx2.df[, 4])
length(slvEx2.df[, 4])
```

```{r ex2_part15_calc_lik}
ex2.expt15.clust1.log.lik = -sum(log(c(5, 11, 13, 14)))
ex2.expt15.clust2.log.lik = - (log(13) + log(14))
ex2.expt15.clust3.log.lik = -sum(log(c(5, 11, 13, 14)))
ex2.expt15.clust4.log.lik = -sum(log(c(6, 13, 14)))

ex2.expt15 = ex2.expt15.clust1.log.lik + 
  ex2.expt15.clust2.log.lik + 
  ex2.expt15.clust3.log.lik + 
  ex2.expt15.clust4.log.lik
ex2.ans15 = log(calcLikIntAllPart(eltsAllPartSet = list(list(c(1), c(2), c(3), c(4))), 
                         sample = slvEx2.df, 
                         alphaC = 3, 
                         betaC = 2))

checkAbsDiff(expected = ex2.expt15, 
             answer = ex2.ans15)
checkAbsRelativeDiff(expected = ex2.expt15,
                     answer = ex2.ans15,
                     rel = min(abs(c(ex2.expt15, ex2.ans15))))
```



```{r ex2_part15_calc_mdp}
ex2.mdp.expt15 = 64/(3 * 17 * 19)
ex2.mdp.ans15 = calcMDPColPrior(alpha = 16, 
                                setList = list(c(1), c(2), c(3), c(4)), 
                                setCountMax = 4)

checkAbsDiff(expected = ex2.mdp.expt15,
             answer = ex2.mdp.ans15)
checkAbsRelativeDiff(expected = ex2.mdp.expt15,
                     answer = ex2.mdp.ans15,
                     rel = min(abs(c(ex2.mdp.expt15, ex2.mdp.ans15))))
```

## Test \texttt{calcLogSubTypeLik} Function

```{r ex2_subtype_lik}
ex2.subtype.lik.expt = ex2.mdp.expt1*exp(ex2.expt1) +
  ex2.mdp.expt2*exp(ex2.expt2) +
  ex2.mdp.expt3*exp(ex2.expt3) +
  ex2.mdp.expt4*exp(ex2.expt4) +
  ex2.mdp.expt5*exp(ex2.expt5) +
  ex2.mdp.expt6*exp(ex2.expt6) +
  ex2.mdp.expt7*exp(ex2.expt7) +
  ex2.mdp.expt8*exp(ex2.expt8) +
  ex2.mdp.expt9*exp(ex2.expt9) +
   ex2.mdp.expt10*exp(ex2.expt10) +
   ex2.mdp.expt11*exp(ex2.expt11) +
   ex2.mdp.expt12*exp(ex2.expt12) +
   ex2.mdp.expt13*exp(ex2.expt13) +
   ex2.mdp.expt14*exp(ex2.expt14) +
   ex2.mdp.expt15*exp(ex2.expt15)

ex2.subtype.log.lik.expt = log(ex2.subtype.lik.expt)

elts4AllPartSetMDPPrior = unlist(lapply(elts4AllPartSet, 
                                        calcMDPColPrior, 
                                        alpha = 16, 
                                        setCountMax = 4))
ex2.subtype.log.lik.ans = 
  calcLogSubTypeLik(eltsAllPartSetList = list(elts4AllPartSet),
                    eltsAllPartSetPriorList = list(elts4AllPartSetMDPPrior),
                    sample = slvEx2.df,
                    mkrList = list(colnames(slvEx2.df)),
                    alphaC = 3,
                    betaC = 2)

checkAbsDiff(expected = ex2.subtype.log.lik.expt, 
             answer = ex2.subtype.log.lik.ans)
checkAbsRelativeDiff(expected = ex1.subtype.log.lik.expt,
                     answer = ex1.subtype.log.lik.ans, 
                     rel = min(abs(c(ex1.subtype.log.lik.expt, 
                                     ex1.subtype.log.lik.ans))))
```

# Test \texttt{calcLogSubTypeLik} Function

```{r ex_subtype_lik}
ex.subtype.log.lik.expt = 
  ex1.subtype.log.lik.expt + ex2.subtype.log.lik.expt
ex.subtype.log.lik.ans = 
  calcLogSubTypeLik(eltsAllPartSetList = list(elts3AllPartSet,
                                              elts4AllPartSet),
                    eltsAllPartSetPriorList = list(elts3AllPartSetMDPPrior,
                                                   elts4AllPartSetMDPPrior),
                    sample = cbind(slvEx1.df, slvEx2.df),
                    mkrList = list(colnames(slvEx1.df), colnames(slvEx2.df)),
                    alphaC = c(2, 3),
                    betaC = c(3, 2))

checkAbsDiff(expected = ex.subtype.log.lik.expt, 
             answer = ex.subtype.log.lik.ans)
checkAbsRelativeDiff(expected = ex.subtype.log.lik.expt,
                     answer = ex.subtype.log.lik.ans, 
                     rel = min(abs(c(ex.subtype.log.lik.expt, 
                                     ex.subtype.log.lik.ans))))
```

# Test \texttt{calcLogTypeLik} Function

```{r ex_type_lik, eval = F}

subtype_lik_test1 = c()
subtype_lik_test2 = c()
ex.type.log.lik.expt = c()
ex.type.log.lik.ans = c()
for(partIndex in 1:length(allPartSets10List)){
  
  ex.subtype.log.lik.expt = c()
  
  for(setIndex in 1:length(allPartSets10List[[partIndex]])){
    obsPartIndexes = allPartSets10List[[partIndex]][[setIndex]]
    ex.subtype.log.lik.expt[setIndex] = 
      calcLogSubTypeLik(eltsAllPartSetList = list(elts3AllPartSet,
                                                  elts4AllPartSet),
                        eltsAllPartSetPriorList = list(elts3AllPartSetMDPPrior,
                                                       elts4AllPartSetMDPPrior),
                        sample = cbind(slvEx1.df[obsPartIndexes, ], 
                                       slvEx2.df[obsPartIndexes, ]),
                        mkrList = list(colnames(slvEx1.df), colnames(slvEx2.df)),
                        alphaC = c(2, 3),
                        betaC = c(3, 2))
  }
  ex.type.log.lik.expt[partIndex] = sum(ex.subtype.log.lik.expt)
  
  setIndexes = sample(1:10, length(allPartSets10List[[partIndex]]))
  obsPartSets = list()
  for(index in 1:length(setIndexes)){
    obsPartSets[[setIndexes[index]]] = allPartSets10List[[partIndex]][[index]]
  }
  
  ex.type.log.lik.ans[partIndex] = calcLogTypeLik(eltsAllPartSetList = list(elts3AllPartSet,
                                                                            elts4AllPartSet), 
                                       eltsAllPartSetPriorList = list(elts3AllPartSetMDPPrior,
                                                                      elts4AllPartSetMDPPrior), 
                                       sample = cbind(slvEx1.df, slvEx2.df), 
                                       subtypeList = obsPartSets, 
                                       mkrList = list(colnames(slvEx1.df), colnames(slvEx2.df)), 
                                       alphaC = c(2, 3), 
                                       betaC = c(3, 2))
  
  subtype_lik_test1[partIndex] = checkAbsDiff(expected = ex.type.log.lik.expt[partIndex], 
                                              answer = ex.type.log.lik.ans[partIndex],
                                              print = FALSE)
  subtype_lik_test2[partIndex] = checkAbsRelativeDiff(expected = ex.type.log.lik.expt[partIndex],
                                                      answer = ex.type.log.lik.ans[partIndex], 
                                                      rel = min(abs(c(ex.subtype.log.lik.expt[partIndex], 
                                                                      ex.subtype.log.lik.ans[partIndex]))),
                                                      print = FALSE)
  if(partIndex %%1000 ==0){
    message(partIndex)
  }
}
```

```{r save, eval = F, echo = F}
ex.type.log.res.df = data.frame(expected = ex.type.log.lik.expt, 
                                answer = ex.type.log.lik.ans)
write.csv(ex.type.log.res.df,
          file = "/Users/chwu/Documents/research/bfc/github/Forensic-Fluids/output/ex.type.log.res.csv",
          quote = F, row.names = F)

save(ex.type.log.lik.expt,
     ex.type.log.lik.ans,
     subtype_lik_test1,
     subtype_lik_test2,
     file = "/Users/chwu/Documents/research/bfc/github/Forensic-Fluids/output/ex.type.log.res.rda")
```
